---
title: "1. part-to-whole"
author: "S.Y. Husada"
date: "2024-04-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

Compare and visualize Bayern Leverkusen shooting stats in the bundesliga
up to 01 April 2024.

# Preparation

## Library

```{r}
library(tidyverse)
library(dlookr)
library(janitor)
library(worldfootballR)
library(camcorder)
```

## Load Data

Load data from worldfootballR.

```{r}
# dat <-
#   fb_season_team_stats(
#     country = "GER",
#     gender = "M",
#     season_end_year = "2024", tier = "1st", stat_type = "shooting")
```

Save and load data in and from RDS.

```{r}
# saveRDS(object = dat, file = "b2324s.RDS")
bundes_23_24_shooting <- readRDS("b2324s.RDS")
```

```{r}
bundes_23_24_shooting %>% head()
```

## Clean Data

```{r}
bundes_23_24_shooting <- bundes_23_24_shooting %>% clean_names() 
```

```{r}
bundes_23_24_shooting %>% head()
```

```{r}
bundes_23_24_shooting %>% get_dupes()
```

```{r}
anyNA(bundes_23_24_shooting)
```

```{r}
bundes_23_24_shooting_team  <- bundes_23_24_shooting %>% 
  filter(team_or_opponent == "team") %>% 
  remove_constant() %>% 
  mutate(squad = as_factor(squad))
bundes_23_24_shooting_team %>% head()
```

```{r}
bundes_23_24_shooting_leverkusen <- 
  bundes_23_24_shooting_team %>% 
  filter(squad == "Leverkusen")
bundes_23_24_shooting_leverkusen
```

# Understand Data

The data is split into standard and expected category, and each category
are further split into accumulated data and their per90 version.

## Glossary {.tabset}

Pl -- \# of Players Number of Players used in Games

### Standard

#### Accumulative

-   **gls_standard**

```         
Goals scored or allowed
```

-   **sh_standard**

```         
Shots Total Does not include penalty kicks
```

-   **so_t_standard**

```         
Shots on Target Note: Shots on target do not include penalty kicks
```

-   **so_t_percent_standard**

```         
Percentage of shots that are on target Minimum .395 shots per squad game to qualify as a leader Note: Shots on target do not include penalty kicks
```

-   **g_per_sh_standard**

```         
Goals per shot Minimum .395 shots per squad game to qualify as a leader
```

-   **g_per_so_t_standard**

```         
Goals per shot on target Minimum .111 shots on target per squad game to qualify as a leader Note: Shots on target do not include penalty kicks
```

-   **dist_standard**

```         
Average distance, in yards, from goal of all shots taken Minimum .395 shots per squad game to qualify as a leader Does not include penalty kicks
```

-   **fk_standard**

```         
Shots from Free Kicks
```

-   **pk_standard**

```         
Penalty Kicks Made
```

-   **p_katt_standard**

```         
Penalty Kicks Attempted
```

#### Per90

-   **sh_per_90_standard**

```         
Shots total per 90 minutes Minimum 30 minutes played per squad game to qualify as a leader
```

-   **so_t_per_90_standard**

```         
Shots on target per 90 minutes Minimum 30 minutes played per squad game to qualify as a leader Note: Shots on target do not include penalty kicks
```

### Expected

-   **x_g_expected**

```         
Expected Goals xG totals include penalty kicks, but do not include penalty shootouts (unless otherwise noted). Provided by Opta. An underline indicates there is a match that is missing data, but will be updated when available.
```

-   **px_g_expected**

```         
Non-Penalty Expected Goals Provided by Opta. An underline indicates there is a match that is missing data, but will be updated when available.
```

-   **npx_g_per_sh_expected**

```         
Non-Penalty Expected Goals per shot Provided by Opta. An underline indicates there is a match that is missing data, but will be updated when available. Minimum .395 shots per squad game to qualify as a leader
```

-   **g_minus_x_g_expected**

```         
Goals minus Expected Goals xG totals include penalty kicks, but do not include penalty shootouts (unless otherwise noted). Provided by Opta. An underline indicates there is a match that is missing data, but will be updated when available.
```

-   **np_g_minus_x_g_expected**

```         
Non-Penalty Goals minus Non-Penalty Expected Goals xG totals include penalty kicks, but do not include penalty shootouts (unless otherwise noted). Provided by Opta. An underline indicates there is a match that is missing data, but will be updated when available.
```

## Descriptive

```{r}
bundes_23_24_shooting_team %>% summary()
```
```{r}
bundes_23_24_shooting_team %>% select(-squad) %>% diagnose_numeric()
```
```{r}
bundes_23_24_shooting_team %>% select(-squad) %>% describe()
```

```{r}
bundes_23_24_shooting_normality <- 
  bundes_23_24_shooting_team %>% 
  select(-squad) %>% 
  normality() %>% 
  arrange(p_value) %>% 
  mutate(is_normal = ifelse(p_value < 0.05, "not_normal", "normal"))

bundes_23_24_shooting_normality
```

#### Deciding Metrics 

I'll use metrics with normal distribution and expected category.

```{r}
bundes_23_24_shooting_normality %>% 
  filter(
    is_normal == "normal"
    )
  
```

Metrics I choose are: 
- to measure open play shooting deadliness: 
  np_g_minus_x_g_expected  
- to measure shot from set-piece deadliness: 
  gls_standard / (fk_standard + pk_standard)
- to measure shooting accuracy:
  so_t_percent_standard
- shoting distance:
  dist_standard
  
# Viz


```{r}
b23224st_longer <- 
  bundes_23_24_shooting_team %>% 
  select(np_g_minus_x_g_expected, gls_standard, fk_standard,  pk_standard, so_t_percent_standard, dist_standard) %>% 
  transmute(
    "Open Play Shot" = np_g_minus_x_g_expected,
    "Set-piece Shot" = (gls_standard / (fk_standard + pk_standard)),
    "Accuracy" = so_t_percent_standard,
    "Distance" = dist_standard
  ) %>% 
  pivot_longer(cols = everything(), names_to = "vars", values_to = "value") %>% 
  group_by(vars) %>% 
  summarise_all(mean)

b23224st_longer
```

```{r}
b23224st_longer_leverkusen <- 
  bundes_23_24_shooting_team %>% 
  filter(squad == "Leverkusen") %>% 
  select(np_g_minus_x_g_expected, gls_standard, fk_standard,  pk_standard, so_t_percent_standard, dist_standard) %>% 
  transmute(
    "Open Play Shot" = np_g_minus_x_g_expected,
    "Set-piece Shot" = (gls_standard / (fk_standard + pk_standard)),
    "Accuracy" = so_t_percent_standard,
    "Distance" = dist_standard
  ) %>% 
  pivot_longer(cols = everything(), names_to = "vars", values_to = "value")
b23224st_longer_leverkusen
```


```{r}

b23224st_longer_scaled <- b23224st_longer %>% 
  select(-value) %>% 
  bind_cols(b23224st_longer$value %>% scale(center = F) %>% as_tibble_col()) %>% 
  mutate( value = round(value, 5))
b23224st_longer_scaled
```

```{r}
leverkusen_longer_scaled <- b23224st_longer_leverkusen %>% 
  select(-value) %>% 
  bind_cols(b23224st_longer_leverkusen$value %>% scale(center = F) %>% as_tibble_col()) %>% 
  mutate( value = round(value, 5))
leverkusen_longer_scaled
```



```{r}
gg_record(
  device = "png", 
  width = 1080 * 2, 
  height = 1080 * 2, 
  units = "px", 
  dpi = 320)
```


```{r echo=FALSE}


theme_setting <- 
  theme(
        panel.background = element_rect(fill = "#F3E500"),
        plot.background = element_rect(fill = "#F3E500",colour = "#F3E500"),
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        axis.text = element_text(size = 30,face = "bold"), 
        plot.title = element_text(
          face = "bold",
          size = 51,
          margin = margin(t = 30, b = 5)
          ),
        plot.caption = element_text(
          face = "bold",
          size = 30,
          margin = margin(b = 0,t = 0)
          ),
        plot.subtitle = element_text(
          face = "italic",
          size = 40,
          margin = margin(b = 0,t = 0)
          ),
        plot.title.position = "panel",)

ggplot(data = b23224st_longer_scaled)+
  geom_col(
    aes(x = vars, y = value), 
    fill = "#000000",
    show.legend = TRUE,
    alpha = .8,
    position = position_jitter(seed = 1,width = 0.15)
    )+
  geom_col(
    data = leverkusen_longer_scaled,
    aes(
      x = vars, 
      y = value
    ),
    fill = "#E32221",
    show.legend = TRUE,
    alpha = .6
           )+
  coord_polar()+
  theme_setting+
  labs(title = "Shooting Stats: Leverkusen vs Bundesliga Average",
       subtitle = "by 1 April 2024",
       caption = "Leverkusen (Red) vs Bundesliga Average (Black)")

```

