---
title: "day_2 - neo"
author: "Sasmito Yudha Husada"
format: html
---


```{r}
library(tidyverse)
library(rvest)
```
# Objective

Visualize the progression of Indonesian National Football Team

# Preparation

## Library

```{r}
library(tidyverse)
library(tsibble)
library(rvest)
library()
library(camcorder)
```


## Scrape Data

```{r}
nat_team_wiki <- read_html("https://en.wikipedia.org/wiki/Indonesia_national_football_team")
```

```{r}
table_list <- nat_team_wiki %>% html_elements(".wikitable") %>% html_table()
```

```{r}
aff_df <- table_list[14] %>% as.data.frame()
```

```{r}
first_row_for_names <- aff_df[1,] %>% as_vector() %>% unname() %>% tolower()
first_row_for_names
```
```{r}
colnames(aff_df) <- first_row_for_names
aff_df
```

## Wrangling Data

```{r}

result_level <-
  c("Group stage", "Semi-finals", "Fourth place", "Third place", "Runners-up")

aff_df_cleaned <- aff_df[-1,] %>%
  select(-squad) %>% 
  remove_rownames() %>% 
  filter(year != "Total") %>% 
  mutate(
    year = as.integer(year),
    result = factor(result, levels = result_level),
    position = str_sub(position, start = 1, end = 1) %>% as.integer(),
    pld = as.integer(pld),
    w = as.integer(w),
    d = as.integer(d),
    l = as.integer(l),
    gf = as.integer(gf),
    ga = as.integer(ga)
    )
aff_df_cleaned
```




#### Deciding Metrics

I want to evaluate three rating from these stats:
1. Winning rate of the team, calculated by dividing matches won with matches played.
2. Offensive rate of the team, calculated by dividing goal for with matches played.
3. Defensive rate of the team, calculated by finding the complement of goal against divided by matches played.

```{r}
aff_df_metricized <- 
  aff_df_cleaned %>% 
  mutate(
    win_rate = round(w / pld, 3),
    offensive_rate = round(gf / pld, 3),
    defensive_rate = (1 - round(ga / pld, 3)),
    sty_era = ifelse(year <= 2020, yes = "Before Shin Tae-yong", no = "Shin Tae-yong Era")
  )
aff_df_metricized 
```


```{r}
aff_df_longer <- aff_df_metricized %>% 
  pivot_longer(cols = 10:12, names_to = "num_vars", values_to = "values")
aff_df_longer
```

## Visualization

```{r}
gg_record(
  
  dir = "~/R/30daychartchallenge/my_30daychartchallenge/day_2",
  device = "jpeg", 
  width = 1080 * 2.5, 
  height = 1080 * 1.5, 
  units = "px", 
  dpi = 320
)
```
#### Def Rate

```{r}
theme_setting <- 
  theme(
        panel.background = element_rect(fill = "#ffffff"),
        plot.background = element_rect(fill = "#ffffff",colour = "#ffffff"),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(margin = margin(r = 15, l = 10)), 
        axis.text = element_text(size = 8,face = "bold"), 
        plot.title = element_text(
          face = "bold",
          size = rel(1.2),
          margin = margin(t = 30, b = 5)
          ),
        plot.caption = element_text(
          face = "bold",
          size = rel(0.6),
          margin = margin(b = 0,t = 15)
          ),
        plot.subtitle = element_text(
          face = "italic",
          size = rel(1),
          margin = margin(b = 0,t = 0)
          ),
        plot.title.position = "panel",
        legend.position = "none")
        

defensive_rate <- 
  ggplot(data = aff_df_longer)+
  geom_line(
    data = aff_df_metricized,
    aes(x = year, y = (10 - ga)),
    col = "#ff0000"
    )+
  geom_point(
    shape = 16,
     data = aff_df_metricized,
     aes(
       x = year, 
       y = (10 - ga), 
       size = defensive_rate,
       col = defensive_rate
       ))+
  theme_setting+
  labs(title = "Defensive Rate", y = "10 - Goal Against", x = "Year")+
  scale_color_gradient(low = "black", high = "#ff0000")+
  facet_wrap(vars(sty_era))+
  expand_limits(y = rel(7))

defensive_rate
```
#### Off Rate
```{r}
theme_setting <- 
  theme(
        panel.background = element_rect(fill = "#ffffff"),
        plot.background = element_rect(fill = "#ffffff",colour = "#ffffff"),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(margin = margin(r = 15, l = 10)), 
        axis.text = element_text(size = 6,face = "bold"), 
        plot.title = element_text(
          face = "bold",
          size = rel(1.05),
          margin = margin(t = 30, b = 5)
          ),
        plot.caption = element_text(
          face = "bold",
          size = rel(0.8),
          margin = margin(b = 0,t = 15)
          ),
        plot.subtitle = element_text(
          face = "italic",
          size = rel(1),
          margin = margin(b = 0,t = 0)
          ),
        plot.title.position = "panel",
        legend.position = "none")
        

offensive_rate <- 
  ggplot(data = aff_df_longer)+
  geom_line(
    data = aff_df_metricized,
    aes(x = year, y = gf),
    col = "#ff0000"
    )+
  geom_point(
    shape = 16,
     data = aff_df_metricized,
     aes(
       x = year, 
       y = gf, 
       size = offensive_rate,
       col = offensive_rate
       ))+
  theme_setting+
  labs(title = "Offensive Rate", y = "Goal For", x = "Year", caption = "The brighter and the bigger the dots the better")+
  scale_color_gradient(low = "black", high = "#ff0000")+
  facet_wrap(vars(sty_era))+
  expand_limits(y = rel(7))

offensive_rate
```
#### Win Rate
```{r}
theme_setting <- 
  theme(
        panel.background = element_rect(fill = "#ffffff"),
        plot.background = element_rect(fill = "#ffffff",colour = "#ffffff"),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(margin = margin(r = 15, l = 10)), 
        axis.text = element_text(size = 6,face = "bold"), 
        plot.title = element_text(
          face = "bold",
          size = rel(1.05),
          margin = margin(t = 30, b = 5)
          ),
        plot.caption = element_text(
          face = "bold",
          size = rel(0.8),
          margin = margin(b = 0,t = 10)
          ),
        plot.subtitle = element_text(
          face = "italic",
          size = rel(1),
          margin = margin(b = 0,t = 0)
          ),
        plot.title.position = "panel",
        legend.position = "none")
        

win_rate <- 
  ggplot(data = aff_df_longer)+
  geom_line(
    data = aff_df_metricized,
    aes(x = year, y = w),
    col = "#ff0000"
    )+
  geom_point(
    shape = 16,
     data = aff_df_metricized,
     aes(
       x = year, 
       y = w, 
       size = win_rate,
       col = win_rate
       ))+
  theme_setting+
  labs(title = "Win Rate", y = "Matches Won", x = "Year")+
  scale_color_gradient(low = "black", high = "#ff0000")+
  facet_wrap(vars(sty_era))+
  expand_limits(y = rel(7))

win_rate
```

#### Cowplot
```{r}
title <- ggdraw() + 
  draw_label(
    "Progression of Indonesian National Team",
    fontface = 'bold'
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(5, 0, 0, 10)
  )

plot_grid(
  title, 
  defensive_rate, 
  offensive_rate, 
  ncol = 1, 
  rel_heights = c(0.5, 3.5, 4), 
  theme = theme(plot.background = element_rect(fill = "white"))
)
```

