---
title: "day_3 - makeover"
author: "Sasmito Y. H."
format: html
---


# Objective

I want to visualize the changes happened at Bayer Leverkusen before their incredible improvement this season. Let's focus on the player transfers happened under the management of Xabi Alonso to narrow the scope.

Xabi Alonso was hired at the 5th of October in 2022, while Bundesliga usually starts at early Augus and ends at late May. Considering the transfer windows happens in June to August (Summer) and in January (Winter), Alonso didn't have an influence in 2022 summer window. So to evaluate his interference, we can only consider:
- Winter transfer window in 22/23 season, 
- Summer transfer window in 23/24 season,
- Winter transfer window in 23/24 season

# Prepare

## Library

```{r}
library(worldfootballR)
library(tidyverse)
```

## Load Data

### Transfer Data

```{r}
# bundesliga_url_2223 <- tm_league_team_urls(country_name = "Germany", start_year = 2022)
```

```{r}
# bundesliga_url_2324 <- tm_league_team_urls(country_name = "Germany", start_year = 2023)
```

```{r}
# bundesliga_tf_2223 <-
#   tm_team_transfers(
#     team_url = bundesliga_url_2223,
#     transfer_window = "winter")
```
```{r}
# bundesliga_tf_2324 <-
#   tm_team_transfers(
#     team_url = bundesliga_url_2324,
#     transfer_window = "all")
```
```{r}
# saveRDS(object = bundesliga_tf_2223, file = "bundesliga_tf_2223.RDS")
# saveRDS(object = bundesliga_tf_2324, file = "bundesliga_tf_2324.RDS")
```

```{r}
# bundesliga_tf_2223 <- readRDS(file = "bundesliga_tf_2223.RDS")
# bundesliga_tf_2324 <- readRDS(file = "bundesliga_tf_2324.RDS")
```

```{r}
leverkusen_tf_2223 <- 
  bundesliga_tf_2223 %>% 
  filter(team_name == "Bayer 04 Leverkusen")
```

```{r}
leverkusen_tf_2324 <- 
  bundesliga_tf_2324 %>% 
  filter(team_name == "Bayer 04 Leverkusen")
```

```{r}
leverkusen_tf_xabi <-
  bind_rows(leverkusen_tf_2223, leverkusen_tf_2324)
```

### Squad Data

```{r}
# bundesliga_league_table_2223_2324 <-
#   fb_season_team_stats(
#     country = "GER", 
#     gender = "M", 
#     season_end_year = c(2023, 2024),
#     tier = "1st",
#     stat_type = "league_table")
```
```{r}
# saveRDS(
#   object = bundesliga_league_table_2223_2324, 
#   file = "bundesliga_league_table_2223_2324.RDS")
```

```{r}
# bundesliga_league_table_2223_2324 <-
#   readRDS("bundesliga_league_table_2223_2324.RDS")
```

### Modify Data

```{r}
lvk_4viz_df <-
leverkusen_tf_xabi %>% 
  filter(transfer_type %in% c("Arrivals", "Departures")) %>% 
  arrange(desc(transfer_fee)) %>%
  head(n = 15) %>% 
  mutate(tf_direction = ifelse(
    transfer_type == "Departures", transfer_fee * -1, transfer_fee
  )) %>% 
  mutate(
    transfer_fee_label = scales::number(
      transfer_fee, scale = 1/1e6, suffix = " M")
  )
```

```{r}
leverkusen_table <-
  bundesliga_league_table_2223_2324 %>% 
  filter(Squad == "Leverkusen")
```


# Viz

## Top 15 transfers
```{r}
group_color <- c("#F3E500", "#E32221")
names(group_color) <- c("Arrivals", "Departures")

xt <- "Transfer Fee"
yt <- ""
mt <- "Leverkusen Player Transfers Under Xabi Alonso"


theme_setting <-
  theme_minimal(base_size = 10)+ 
  theme(
    plot.title.position = "panel", 
    plot.subtitle = ggtext::element_markdown(),
    title = element_text(size = rel(1.5)),
    legend.text = element_text(size = rel(1.3))
    )

ggplot(lvk_4viz_df,
    aes(
      x = tf_direction,
      y = reorder(player_name, transfer_fee),
      fill = transfer_type
    )
  ) +
  geom_col(alpha = 0.45) +
  geom_text(
    size = 2.8,
    aes(
      label = player_name, 
      hjust = ifelse(tf_direction > 0, 1.05, -0.05)
      )
    )+
  geom_text(
    size = 3,
    aes(
      label = transfer_fee_label, 
      hjust = ifelse(tf_direction > 0, -0.25, 1.2)
      )
    )+
  theme_setting+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
        
        )+
  labs(y = yt, title = mt, x = xt, fill = "Team")+
  coord_cartesian(
    xlim = 
      c(
        min(lvk_4viz_df$tf_direction)-1.2*mean(lvk_4viz_df$transfer_fee),
        max(lvk_4viz_df$tf_direction)+1.2*mean(lvk_4viz_df$transfer_fee))
  )+
  scale_fill_manual(values = group_color)
```
## Table Comparison
```{r}
theme_setting <-
  theme_minimal(base_size = 10)+ 
  theme(
    plot.title.position = "panel", 
    plot.subtitle = ggtext::element_markdown(),
    title = element_text(size = rel(1.5)),
    legend.text = element_text(size = rel(1.3)),
    plot.margin = margin(20, 30, 40, 30),
    axis.text.x = element_blank()
    )

leverkusen_table %>% 
  ggplot(
    aes(x = Season_End_Year, y = Pts.MP)
  )+
  geom_line(show.legend = F)+
  geom_point(aes(size = Pts, col = MP), show.legend = F)+
  geom_label(
    data = leverkusen_table %>% filter(Season_End_Year == 2023),
    size = 2.1,
    aes(
      x = Season_End_Year, 
      y = Pts.MP, 
      label = paste0(
        "Season End: ", Season_End_Year, 
        "\nMatch Played: ", MP, 
        "\nPoints Gained: ", Pts)),nudge_y = 0.35)+
  geom_label(
    data = leverkusen_table %>% filter(Season_End_Year == 2024),
    size = 2.1,
    aes(
      x = Season_End_Year, 
      y = Pts.MP, 
      label = paste0(
        "Season Ongoing: ", Season_End_Year, 
        "\nMatch Played: ", MP, 
        "\nPoints Gained: ", Pts)),nudge_y = -0.35
  )+
  labs(
    title = "Leverkusen Points per Match Played", 
    subtitle = "22/23 vs 23/24",
    x = element_blank(),
    y = "Points per Match"
    )+
  scale_x_continuous(breaks = c(2023, 2024))+
  xlim(2023 - 0.25, 2024 + 0.25)+
  ylim(1.25, 3.25)+
  theme_setting+
  scale_color_gradient(low =  "#E32221", high = "#F3E500" )
  
```

